New insights into the Weddell Sea ecosystem applying a network approach

### 2022-11-17 Manuscript/WordFormat.Rmd
Rmarkdown for word format

---
title: New insights into the Weddell Sea ecosystem applying a network approach
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Tomás I.
    surname: Marina
    affiliation: "1, *"
    email: tomasimarina@gmail.com
  - given_name: Leonardo A.
    surname: Saravia
    affiliation: "1, *"
    email: arysar@gmail.com
  - given_name: Susanne
    surname: Kortsch
    affiliation: 2
    # authors can have multiple affiliations, which can also be used to mark deceased coauthors
# If you have more than one corresponding author, add them manually using the following structure (note the commas):
# Two authors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# Three authors or more: Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry (j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)
# If the following line is uncommented, the "corresponding: true" above are ignored
correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: University of Helsinki, Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  The network approach has proved to shade light on the structure and stability of complex marine ecosystems. In recent years, such approach was succesfully applied to the study of polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea, particularly the high continental shelf area (74-78ºS), is one of the most studied areas outside the Antarctic Peninsula in the Southern Ocean. However, few researches have considered the known complexity of the Weddell Sea. Understanding and projecting impacts of change in this ecosystem requires a framework considering the full complexity of it. Here we focus on the trophic interactions that underpin ecosystem structure and stability. We analysed the Weddell Sea food web, composed by 490 species and 16041 predator-prey interactions, which is the highest-resolved Antarctic food web constructed so far. The objective of our work was threefold: 1) estimate the strength for each interaction in the food web, 2) characterise the species considering weighted and unweighted properties, and 3) analyse the species' role in the stability of the food web. On the one hand, we found that the distribution of the interaction strength at the food web level is asymmetric, where weak interactions are prevalent. Including such information as a (weighted) property for species we detected a positive relationship between species mean interaction strength and two unweighted properties, trophic level and the total number of interactions. We also found that only few species are key in terms of food web stability, presenting particular weighted and unweighted properties. In the same analysis we have integrated food web and species information, enabling a more complete assessment of the ecosystem structure and function, likely highlighting the ecological processes at play in the Weddell Sea. We consider that our results provide new insights important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

bibliography: ../WeddellSea.bib
running:
  title: New insights into the Weddell Sea food web
  author: Marina et al.
competinginterests: |
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
# copyrightstatement: |
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #  use this to add a statement when having only software code available
  #  use this to add a statement when having only data sets available
  #code: |
  #data: |
codedata: |
    R code and data set used in this work is available at https://github.com/EcoComplex/WeddellSea
#   sample: |
# videosupplement: |
authorcontribution: |
  TIM and LAS: Conceptualization (lead); Data curation (lead); Formal analysis (lead); Methodology (lead); Coding (lead); Writing – original draft (lead); Writing – review and editing (lead). SK: Conceptualization (lead); Formal analysis (supporting); Methodology (supporting); Coding (supporting); Writing – original draft (supporting); Writing – review and editing (supporting).
# disclaimer: |
  # We like Copernicus.
acknowledgements: |
  Thanks to the rticles contributors!
# appendix: |
#   \section{Figures and tables in appendices}
#   \subsection{Option 1}
#   If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
#   They will be correctly named automatically.
#   \subsection{Option 2}
#   If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
#   
#   `\appendixfigures` needs to be added in front of appendix figures
#   `\appendixtables` needs to be added in front of appendix tables
#   
#   Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
#   Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
#   To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  rticles::copernicus_article: 
    highlight: NULL
    keep_tex: true
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
editor_options: 
  markdown: 
    wrap: 72
---

\introduction[Introduction]

@@

# Methodology

## Study area

@@

## Weddell Sea food web dataset

@@

## Dataset analyses

@@

@@

# Results

## Interaction strength

@@

## Species properties

@@

## Extinction simulations and stability

@@

@@

# Discussion

@@

\clearpage
\conclusions[Conclusions]

@@

\clearpage

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Map of the Weddell Sea and Dronning Maud Land sector highlighting the high Antarctic shelf as a dashed-line contour. Modified from www.soos.aq."}
knitr::include_graphics("Fig.1_StudyMap.jpg")
```

\clearpage

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Scheme of a network showing the weighted and unweighted properties we used to characterize the species of the Weddell Sea food web."}
knitr::include_graphics("Fig.2_ToyFoodWeb.jpg")
```

\clearpage

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Frequency distribution of interaction strengths for the Weddell Sea food web. Total number of interactions = 16041."}
knitr::include_graphics("Fig3_IntDist.png")
```

\clearpage

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Relationships between weighted (mean Interaction Strength) and unweighted properties including habitat. Linear regressions are shown between log(mean interaction strength) and trophic level (A), degree (B) and trophic similarity (C). Linear regressions for trophic level ($y = 1.12x - 15.29, R^2 = 0.43, p-value < 2e-16$), degree ($y = 0.006x - 12.77, R^2 = 0.03, p-value = 4.06e-5$) and trophic similarity ($y = -1.46x - 12.18, R^2 = -0.0004, p-value = 0.36$)."}
knitr::include_graphics("Fig4_LinReg.png")
```

\clearpage

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Quasi-Sign Stability (QSS) difference between the whole Weddell Sea food web (n = 490) and the food web without one species (n = 489) for weighted (interaction strength) and unweighted species properties, and habitat. Point color indicates the impact on the QSS; if significant the extinction of that species altered the stability (QSS) of the food web."}
knitr::include_graphics("Fig.5_QSSDif.png")
```

\clearpage

\clearpage

```{=tex}
\begin{table}[t]
\caption{Properties of the species that when become extinct generated a significant impact on the stability of the Weddell Sea food web, ordered by significance (Anderson-Darling p-value). References: meanIS = mean interaction strength, TL = trophic level, Deg = degree, TS = trophic similarity, StabDif = stability difference, ADvalue = Anderson-Darling p-value.}
\begin{tabular}{l c c c c c c c}
\tophline

\textbf{Species} & \textbf{meanIS} & \textbf{TL} & \textbf{Deg} & \textbf{TS} & \textbf{Habitat} & \textbf{StabDif} & \textbf{ADvalue}\\
\middlehline
Orcinus orca & 1.83e-4 & 5.03 & 26 & 0.037 & Benthopelagic & 4.67e-5 & 2.28e-41 \\
\middlehline
Macrourus holotrachys & 8.30e-5 & 4.70 & 85 & 0.112 & Benthopelagic & 3.55e-5 & 2.73e-23 \\
\middlehline
Pagetopsis macropterus & 7.08e-5 & 4.64 & 76 & 0.113 & Demersal & -1.80e-5 & 2.38e-12 \\
\middlehline
Abyssorchomene nodimanus & 2.56e-5 & 4.21 & 137 & 0.130 & Benthopelagic & 2.30e-5 & 8.52e-10 \\
\middlehline
Dissostichus mawsoni & 7.82e-5 & 4.12 & 87 & 0.126 & Pelagic & 2.17e-5 & 1.57e-9 \\
\middlehline
Macrourus whitsoni & 7.14e-5 & 4.55 & 92 & 0.124 & Benthopelagic & 2.12e-5 & 3.30e-8 \\
\middlehline
Hydrurga leptonyx & 1.03e-4 & 4.72 & 67 & 0.094 & Land-based & 2.04e-5 & 9.66e-6 \\
\middlehline
Mesonychoteuthis hamiltoni & 1.80e-4 & 4.41 & 29 & 0.028 & Pelagic & 1.82e-5 & 4.59e-5 \\
\middlehline
Champsocephalus gunnari & 7.62e-5 & 3.72 & 46 & 0.086 & Pelagic & 1.83e-5 & 6.79e-5 \\
\middlehline
Notothenia marmorata & 8.27e-5 & 4.09 & 44 & 0.091 & Demersal & 1.60e-5 & 1.23e-4 \\
\middlehline
Arctocephalus gazella & 9.28e-5 & 4.67 & 61 & 0.093 & Land-based & 1.17e-5 & 2.09e-4 \\
\middlehline
Trematomus pennellii & 3.04e-5 & 4.04 & 192 & 0.158 & Demersal & 1.44e-5 & 1.00e-3 \\
\middlehline
Mirounga leonina & 1.20e-4 & 4.87 & 56 & 0.080 & Land-based & 1.41e-5 & 1.28e-3 \\
\middlehline
Notothenia coriiceps & 4.94e-5 & 4.27 & 130 & 0.126 & Demersal & 1.44e-5 & 1.66e-3 \\
\middlehline
Maxilliphimedia longipes & 2.21e-6 & 3.26 & 60 & 0.136 & Benthopelagic & -4.46e-6 & 9.74e-3 \\

\appendixfigures
\clearpage

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Graphic representation of the Weddell Sea food web. Species (nodes) are arranged vertically and colored by trophic level. The diameter of the node indicates the total number of interactions. Predator-prey interactions are represented by the arrows, from prey to predator."}
knitr::include_graphics("App1_FWplot.png")
```

>>>>

---
title: "New insights into the Weddell Sea ecosystem applying a network approach"
author: "Tomás I. Marina, Leonardo A. Saravia and Susanne Kortsch"
bibliography: ../WeddellSea.bib
output: word_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Abstract

The network approach has proved to shade light on the structure and stability of complex marine ecosystems. In recent years, such approach was succesfully applied to the study of polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea, particularly the high continental shelf area (74-78ºS), is one of the most studied areas outside the Antarctic Peninsula in the Southern Ocean. However, few researches have considered the known complexity of the Weddell Sea. Understanding and projecting impacts of change in this ecosystem requires a framework considering the full complexity of it. Here we focus on the trophic interactions that underpin ecosystem structure and stability. We analysed the Weddell Sea food web, composed by 490 species and 16041 predator-prey interactions, which is the highest-resolved Antarctic food web constructed so far. The objective of our work was threefold: 1) estimate the strength for each interaction in the food web, 2) characterise the species considering weighted and unweighted properties, and 3) analyse the species' role in the stability of the food web. On the one hand, we found that the distribution of the interaction strength at the food web level is asymmetric, where weak interactions are prevalent. Including such information as a (weighted) property for species we detected a positive relationship between species mean interaction strength and two unweighted properties, trophic level and the total number of interactions. We also found that only few species are key in terms of food web stability, presenting particular weighted and unweighted properties. In the same analysis we have integrated food web and species information, enabling a more complete assessment of the ecosystem structure and function, likely highlighting the ecological processes at play in the Weddell Sea. We consider that our results provide new insights important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

## Introduction

@@

## Methodology

### Study area

@@

```{r echo=FALSE, fig.align='center', fig.cap="Figure 1. Map of the Weddell Sea and Dronning Maud Land sector highlighting the high Antarctic shelf as a dashed-line contour. Modified from www.soos.aq."}

library(knitr)
knitr::include_graphics("Fig.1_StudyMap.jpg", dpi = 600)

```

### Weddell Sea food web dataset

@@

### Dataset analyses

@@

```{r echo=FALSE, fig.align='center', fig.cap="Figure 2. Scheme of a network showing the weighted and unweighted properties we used to characterize the species of the Weddell Sea food web."}

library(knitr)
knitr::include_graphics("Fig.2_ToyFoodWeb.jpg", dpi = 600)

```

@@

## Results

### Interaction strength

@@

```{r echo=FALSE, fig.align='center', fig.cap="Figure 3. Frequency distribution of interaction strengths for the Weddell Sea food web. Total number of interactions = 16041."}

library(knitr)
knitr::include_graphics("Fig3_IntDist.png", dpi = 600)

```

```{r echo=FALSE, results='asis'}
library(knitr)
library(dplyr)

load("../Manuscript/ModelFit.rda")
knitr::kable(IS_fit %>% arrange(., deltaAIC), digits = c(0,0,2,2), col.names = c("Model", "df", "AIC", "deltaAIC"), caption = "Model comparison for the distribution of interaction strengths of the Weddell Sea food web. Order by best fit. References: df = degrees of freedom, AIC = Akaike Information Criterion, deltaAIC = difference with best fit. Log-Normal is the best model.")

```

@@

```{r echo=FALSE, fig.align='center', fig.cap="Figure 4. Relationships between weighted (mean Interaction Strength) and unweighted properties including habitat. Linear regressions are shown between log(mean interaction strength) and trophic level (A), degree (B) and trophic similarity (C). Linear regressions for trophic level ($y = 1.12x - 15.29, R^2 = 0.43, p-value < 2e-16$), degree ($y = 0.006x - 12.77, R^2 = 0.03, p-value = 4.06e-5$) and trophic similarity ($y = -1.46x - 12.18, R^2 = -0.0004, p-value = 0.36$)."}

library(knitr)
knitr::include_graphics("Fig4_LinReg.png", dpi = 600)

```

@@

```{r echo=FALSE, fig.align='center', fig.cap="Figure 5. Quasi-Sign Stability (QSS) difference between the whole Weddell Sea food web (n = 490) and the food web without one species (n = 489) for weighted (interaction strength) and unweighted species properties, and habitat. Point color indicates the impact on the QSS; if significant the extinction of that species altered the stability (QSS) of the food web."}

library(knitr)
knitr::include_graphics("Fig.5_QSSDif.png", dpi = 600)

```

@@

```{r echo=FALSE, results='asis'}
library(knitr)
library(dplyr)

load("QSS_summary_sep22.rda")
knitr::kable(all_data %>% dplyr::select(TrophicSpecies, IS_mean, TL, TotalDegree, meanTrophicSimil, Habitat, difQSS, Ad_pvalue) %>% dplyr::filter(Ad_pvalue < 0.01) %>%  arrange(., Ad_pvalue), digits = c(0,7,2,0,3,0,7,41), col.names = c("Species","MeanIS", "TL", "Deg", "TS", "Habitat", "StabDif", "ADvalue"), caption = "Properties of the species that when become extinct generated a significant impact on the stability of the Weddell Sea food web, ordered by significance (Anderson-Darling p-value). References: meanIS = mean interaction strength, TL = trophic level, Deg = degree, TS = trophic similarity, StabDif = stability difference, ADvalue = Anderson-Darling p-value.")

```

## Discussion

@@

## Conclusions

@@

## References
